Nuclei Segmentation in Histopathology images using Deep Neural Networks

### 2016-11-07 Paper.tex
shorter introduction

% Template for ISBI-2016 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[francais]{babel}
\usepackage{hyperref}
\usepackage{seqsplit}

@@

\begin{document}
%\ninept
%
\maketitle
\vspace{-1cm}
%
\begin{abstract}
Analysis and interpretation of stained tumor sections is one of the
main tools in cancer diagnosis and prognosis, which is mainly carried
out manually by pathologists. The avent of digital pathology provides
us with the challenging opportunity to automatically analyze large
amounts of these complex image data in order to draw
biological conclusions from them and to study cellular and tissular
phenotypes at a large scale. One of the bottlenecks for such
approaches is the automatic segmentation of cell nuclei from this type
of image data. Here, we present a fully automated workflow to segment
nuclei from histopathology image data by using deep neural
networks trained from a set of manually annotated images and by processing
the posterior probability maps in order to split jointly segmented
nuclei. Further, we provide the image data set that has been generated
for this study as a benchmark set to the scientific community.  
\end{abstract}
%
\begin{keywords}
Deep Learning, Convolutional Neural Networks, Nuclei Segmentation,
Histopathology, Digital Pathology, Breast Cancer, Cellular Phenotyping
\end{keywords}
%
\section{Introduction}
\label{sec:intro}

@@

% two approaches: CAD vs. understanding, importance of segmentation
For these reasons, analysis of histopathology data has received much
attention over the last years. While many approaches aim at
automatically performing the task of a human pathologist (Computer
Aided Diagnosis, CAD), we want to investigate the influence of
biologically relevant features, such as phenotype distributions or
heterogeneity measures from the tissue data on different clinical
variables, such as resistance to treatment or relapse probability.
One essential step is thus the segmentation of nuclei from tissue
images,  in order to assign them a cell type or phenotypic label and
to evaluate their spatial distribution.

@@

\section{Different architectures/Results}
\label{sec:results}
We experience with 3 known arhitectures in semantic segmentation,
named PangNet, Fully Convolutionnal Net (FCN) and DeconvNet. The most
basic architecture, PangNet, consists of 4  
convolutional layers where each convolutional layer has 8 feature
maps\cite{pang2010cell}.  
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
architecture has the advantage of re-using a classical deep learning 
architecture with additional upsampling and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture\cite{long2015fcn}. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layers as one intends to learn 
the proper upsampling through repeated deconvolution and convolution 
layers. This model can also be fine tuned\cite{noh2015learning}. We also did an
ensemble classifier of these two previous nets as suggested in 
\cite{noh2015learning}.
% The final architecture that we tried is the UNet. This network combines 
%both contribution of the above networks, the UNet has a similar 
%architecture to the DeconvNet with skip layers.
\section{Result and discussion}
\label{sec:result}
\noindent We conducted our experiments on GPUs via the caffe framework, 
\cite{jia2014caffe}. We present our results in table \ref{tab:res} where 
we averaged the metrics over all left out patients. We also added, for 
illustration purposes \ref{fig:prediction}, image predictions and 
associated 
probability maps to evaluate some differences between the classifiers. 
We notice that the simpler net, PangNet seems to have learnt simple 
rules such as color information. As soon as the nuclei are not
homogeneous and dark inside, the method fails ( "holes" in the
cells). On the 
contrary, deeper nets as FCN and DeconvNet have learnt deeper features 
and seem capable of recognizing whole cells. On these probability maps 
we applied our watershed segmentation and we can picture the results in 
the third 
column. By construction, our post-processing 
scheme is very sensitive to noise as we can see in the PangNet 
prediction, many unrelated minimas will lead to highly partitionned cell. 
This phenomenom also appears with the DeconvNet where undesired 
cuts were performed, like for the cell at the bottom left of the image. 
We picked $\lambda = 7$ which seemed to give a nice partitionning of 
the image, this value can be found by cross validation or by simple 
visualization of the results.
\begin{table}
\begin{tabular}{|c|c|c|c|c|}
\hline
  & PangNet & DeconvNet & FCN & Ensemble\\
 \hline
Accuracy  &       0.936 & 0.968 &0.958 & 0.968  \\
IU   &    0.759 &     0.856 & 0.857 & 0.861 \\
Recall     &       0.744  &     0.856 & 0.855 & 0.876\\
Precision   &       0.741 &     0.858 & 0.878 & 0.851\\
F1    &       0.742 &     0.856 & 0.866  & 0.863\\
TP       &       0.744 &     0.856 & 0.855 & 0.876 \\
TN      &       0.963 &     0.981 &  0.977 & 0.979\\
Performance    &       0.853 &     0.918 & 0.916 & 0.928\\
Pixel error  &  0.063 &     0.032 & 0.042 & 0.032\\
\hline
\end{tabular}
\caption{Results}
\label{tab:res}
\end{table}

@@

We have presented a fully automized workflow for segmenting nuclei in
histopathology images based on deep learning and mathematical
morphology. We have also generated a manually curated segmentation
data base, which we make publicly available. We believe that such a
method can become very useful for the investigation of cellular
phenotypes at a tissue level and their relation to disease.

@@

>>>>

% Template for ISBI-2016 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[francais]{babel}
\usepackage{hyperref}

@@

\begin{document}
%\ninept
%
\maketitle

%
\begin{abstract}
Analysis and interpretation of stained tumor sections is one of the
main tools in cancer diagnosis and prognosis, which is mainly carried
out manually by pathologists. The avent of digital pathology provides
us with the challenging opportunity to automatically analyze large
amounts of these complex image data in order to draw
biological conclusions from them and to study cellular and tissular
phenotypes at a large scale. One of the bottlenecks for such
approaches is the automatic segmentation of cell nuclei from this type
of image data. Here, we present a fully automated workflow to segment
nuclei from histopathology image data by using deep neural
networks trained from a set of manually annotated images and by processing
the posterior probability maps in order to split jointly segmented
nuclei. Further, we provide the image data set that has been generated
for this study as a benchmark set to the scientific community.  
\end{abstract}
%
\begin{keywords}
Deep Learning, Convolutional Neural Networks, Nuclei Segmentation,
Histopathology, Digital Pathology, Breast Cancer, Cellular Phenotyping
\end{keywords}
%
\section{Introduction}
\label{sec:intro}

@@

% two approaches: CAD vs. understanding, importance of segmentation
For these reasons, analysis of histopathology data has received much
attention over the last years. In particular for approaches aiming at
relating biologically relevant features (such as phenotype
distributions or heterogeneity measures) to clinical variables,
segmentation of nuclei from tissue images is essential.

%While many approaches aim at
%automatically performing the task of a human pathologist (Computer
%Aided Diagnosis, CAD), we want to investigate the influence of
%biologically relevant features, such as phenotype distributions or
%heterogeneity measures from the tissue data on different clinical
%variables, such as resistance to treatment or relapse probability.
%One essential step is thus the segmentation of nuclei from tissue
%images,  in order to assign them a cell type or phenotypic label and
%to evaluate their spatial distribution.

@@

\section{Different architectures/Results}
\label{sec:results}
\noindent We experience with 3 known arhitectures in semantic segmentation, named PangNet, Fully Convolutionnal Net (FCN) and DeconvNet. The most basic architecture, PangNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map, for more information please refer to \cite{pang2010cell}. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
architecture has the advantage of re-using a classical deep learning 
architecture with additional upsampling and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture\cite{long2015fcn}. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layers as one intends to learn 
the proper upsampling through repeated deconvolution and convolution 
layers. This model can also be fine tuned\cite{noh2015learning}. We also did an
ensemble classifier of these two previous nets as suggested in 
\cite{noh2015learning}.
% The final architecture that we tried is the UNet. This network combines 
%both contribution of the above networks, the UNet has a similar 
%architecture to the DeconvNet with skip layers.
\section{Result and discussion}
\label{sec:result}
\noindent We conducted our experiments on GPUs via the caffe framework, 
\cite{jia2014caffe}. We present our results in table \ref{tab:res} where 
we averaged the metrics over all left out patients. We also added, for 
illustration purposes \ref{fig:prediction}, image predictions and 
associated 
probability maps to evaluate some differences between the classifiers. 
We notice that the simpler net, PangNet seems to have learnt simple 
rules such as color information. As soon as the nuclei are not
homogeneous and dark inside, the method fails ( "holes" in the
cells). On the 
contrary, deeper nets as FCN and DeconvNet have learnt deeper features 
and seem capable of recognizing whole cells. On these probability maps 
we applied our watershed segmentation and we can picture the results in 
the third 
column. By construction, our post-processing 
scheme is very sensitive to noise as we can see in the PangNet 
prediction, many unrelated minimas will lead to highly partitionned cell. 
This phenomenom also appears with the DeconvNet where undesired 
cuts were performed, like for the cell at the bottom left of the image. 
We picked $\lambda = 7$ which seemed to give a nice partitionning of 
the image, this value can be found by cross validation or by simple 
visualization of the results.
\begin{table}
\begin{tabular}{|c|c|c|c|c|}
\hline
  & PangNet & DeconvNet & FCN & Ensemble\\
 \hline
Accuracy  &       0.936 & 0.968 &0.958 & 0.968  \\
IU   &    0.759 &     0.856 & 0.857 & 0.861 \\
Recall     &       0.744  &     0.856 & 0.855 & 0.876\\
Precision   &       0.741 &     0.858 & 0.878 & 0.851\\
F1    &       0.742 &     0.856 & 0.866  & 0.863\\
TP       &       0.744 &     0.856 & 0.855 & 0.876 \\
TN      &       0.963 &     0.981 &  0.977 & 0.979\\
Performance    &       0.853 &     0.918 & 0.916 & 0.928\\
Pixel error  &  0.063 &     0.032 & 0.042 & 0.032\\
\hline
\end{tabular}
\caption{Results}
\label{tab:res}
\end{table}

@@

We presented a fully automized workflow for segmenting nuclei in
histopathology images based on deep learning and mathematical
morphology. We have also generated a manually curated segmentation
data base, which we make publicly available. We believe that such a
method can become very useful for the investigation of cellular
phenotypes at a tissue level and their relation to disease.

@@

%\begin{minipage}[b]{0.32\linewidth}
%  \centering
%  \centerline{\includegraphics[height=2.4cm]{BaochuanB}}
%%  \vspace{1.5cm}
%  \centerline{PangNet}\medskip
%\end{minipage}
%\hfill
%\begin{minipage}[b]{.32\linewidth}
%  \centering
%  \centerline{\includegraphics[height=2.5cm]{BaochuanP}}
%%  \vspace{1.5cm}
%  \centerline{Probability map}\medskip
%\end{minipage}
%\hfill
%\begin{minipage}[b]{.32\linewidth}
%  \centering
%  \centerline{\includegraphics[height=2.4cm]{Baochuan_WS}}
%%  \vspace{1.5cm}
%  \centerline{Watershed}\medskip
%\end{minipage}
%
%%
%%
%\begin{minipage}[b]{0.32\linewidth}
%  \centering
%  \centerline{\includegraphics[height=2.4cm]{DeconvNetB}}
%%  \vspace{1.5cm}
%  \centerline{DeconvNet}\medskip
%\end{minipage}
%\hfill
%\begin{minipage}[b]{.32\linewidth}
%  \centering
%  \centerline{\includegraphics[height=2.4cm]{DeconvNetP}}
%%  \vspace{1.5cm}
%  \centerline{Probability map}\medskip
%\end{minipage}
%\hfill
%\begin{minipage}[b]{.32\linewidth}
%  \centering
%  \centerline{\includegraphics[height=2.4cm]{DeconvNet_WS}}
%%  \vspace{1.5cm}
%  \centerline{Watershed}\medskip
%\end{minipage}
%%
%\begin{minipage}[b]{0.32\linewidth}
%  \centering
%  \centerline{\includegraphics[height=2.4cm]{FCNB}}
%%  \vspace{1.5cm}
%  \centerline{FCN}\medskip
%\end{minipage}
%\hfill
%\begin{minipage}[b]{.32\linewidth}
%  \centering
%  \centerline{\includegraphics[height=2.4cm]{FCNP}}
%%  \vspace{1.5cm}
%  \centerline{Probability map}\medskip
%\end{minipage}
%\hfill
%\begin{minipage}[b]{.32\linewidth}
%  \centering
%  \centerline{\includegraphics[height=2.4cm]{FCN_WS}}
%%  \vspace{1.5cm}
%  \centerline{Watershed}\medskip
%\end{minipage}
%%
%\begin{minipage}[b]{0.32\linewidth}
%  \centering
%  \centerline{\includegraphics[height=2.4cm]{EnsembleB}}
%%  \vspace{1.5cm}
%  \centerline{Ensemble}\medskip
%\end{minipage}
%\hfill
%\begin{minipage}[b]{.32\linewidth}
%  \centering
%  \centerline{\includegraphics[height=2.4cm]{EnsembleP}}
%%  \vspace{1.5cm}
%  \centerline{Probability map}\medskip
%\end{minipage}
%\hfill
%\begin{minipage}[b]{.32\linewidth}
%  \centering
%  \centerline{\includegraphics[height=2.4cm]{Ensemble_WS}}
%%  \vspace{1.5cm}
%  \centerline{Watershed}\medskip
%\end{minipage}
%\caption{Prediction via different classifiers of a random sample on the left out patient: 581910}
%\label{fig:prediction}
%%
%\end{figure}
